First paper - Draft

### 2017-06-01 paper/paper.Rnw
Finished moving script

<<functions_for_modelling, echo = F, cache = TRUE>>=

@@

# Producing the plot to get the difference between the top 30% of the high educated
# vs the bottom 30% of the low educated. This function loops through each dataset/country
# and survey reliability and estimates the difference while also calculating the
# joint standard error of the difference.

@@

results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH")
results_read <- test_diff(adapted_year_data, reliability_pisa, "READ")
# US is missing for reading

@
\end{document}

>>>>

<<functions_for_modelling, echo = F>>=

@@

# Producing the plot to get the difference between the top 30% of the high educated
# vs the bottom 30% of the low educated. This function loops through each dataset/country
# and survey reliability and estimates the difference while also extracting the s.e. of each
# difference.

@@

results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH")
results_read <- test_diff(adapted_year_data, reliability_pisa, "READ")
# US is missing for reading
@

<<ci_for_difference, echo = F>>=
reduced_data <-
    map2(results_math, years, function(.x, .y) {
      .x %>%
        mutate(wave = .y) %>%
        filter(!is.na(escs_dummy))
    }) %>%
    bind_rows() %>%
    mutate(lower = Mean - 1.96 * s.e.,
           upper = Mean + 1.96 * s.e.)
@

<<graphin differences>>=
  reduced_data %>%
    filter(country %in% c("United States")) %>%
    ggplot(aes(as.character(wave), Mean, group = escs_dummy, colour = escs_dummy)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
    geom_line() +
    facet_wrap(~ country)
@

<<>>=
# Calculate the joint standard error of the different
  se_data <-
    reduced_data %>%
    select(country, escs_dummy, wave, s.e.) %>%
    spread(escs_dummy, s.e.) %>%
    transmute(country, wave,
              se_diff = sqrt(abs(`1`^2 - `0`^2)))

<<>>=

# Calculate the different between the gap and together with it's joint s.e graph
# the absolut difference.

countries <- c("Chile", "Austria", "Belgium", "Canada", "Czech Republic", "Denmark",
                 "Finland", "France", "Germany", "Italy", "Japan", "Netherlands", "Norway",
                 "Poland", "Spain", "Sweden", "United Kingdom", "United States")

reduced_data %>%
    select(country, escs_dummy, Mean, wave) %>%
    spread(key = escs_dummy, value = Mean) %>%
    mutate(diff = `1` - `0`) %>%
    gather(escs_dummy, Mean, -country, -wave, -`1`, -`0`) %>%
    left_join(se_data) %>%
    mutate(continent = countrycode(country, "country.name", "continent"),
           region = countrycode(country, "country.name", "region"),
           lower = Mean - 1.96 * se_diff,
           upper = Mean + 1.96 * se_diff) %>%
    filter(!is.na(continent), !is.na(region), country %in% countries) %>%
    ggplot(aes(as.factor(wave), Mean, group = country, colour = country)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
    geom_line() +
    geom_point(size = 0.5) +
    scale_colour_discrete(guide = F) +
    # geom_smooth(aes(group = region), colour = "blue", method = "lm") +
    coord_cartesian(ylim = c(0, 4)) +
    facet_wrap(~ country) +
    ggtitle("math")

# Increase:  
# Sweden - steady increase in both tests
# Austria - increase in math - slight increase in read
# Finland - very sharp increase in both
# France - very sharp increase in both
# Netherlands - sharp increase in both

# Decrease:
# US - decrease in both tests
# Chile - decrease in both tests

# No change:
# Canada - stable red - increase math
# UK - slight decrease red - stable math
# Belgium - no change
# Czech republic - no change
# Denmark no change
# Germany - no change
# ITaly - no change
# Japan -  no change
# Norway - no change
# Poland - no change
# Spain - no change

<<>>=
# Next steps:

# Show country patterns with reading and math together into one graph

# Making country groups of the graphs showing that some countries are increasing, others are steady
# While others are decreasing. Do this with a few countries, but end by showing all countries.

# Continue by studying what's happening in those countries where it's increasing/decreasing
# In the gap graph

# Continue by doing the multilevel models to see what explains what. Include
# all indicators from the reardon/russian girl paper.

# Graph the increase in each country vs the increase/decrease of the economic inequality indicators
# Specially the 90/10

# Calculate how big is the gap between reading and math

# Continue with the PIRLS to see if there are specific patterns in 4th and 8th graders gap.

# Get each country trendline adjusted for the inequality indicators and place in the same country graph.

# Should I add the parent's education in the lm model to see how trends change adjusted for that?

Try one!
\end{document}
